---
title: "Lupus RGMX descriptive analysis"
author: "Hernández-Ledesma Ana Laura"
date: "2023-02-18"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## STEP 1: Loading required packages and reading database
```{r Reading database}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(mxmaps)
library(UpSetR)
data <- read.csv(file = "SLE.csv")
```

## STEP 2: Loading colorblind palette
```{r Colorblind palette definition }
colorblind_palette <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
                        "#44AA99", "#999933", "#882255", "#661100", "#6699CC", "#888888")
scales::show_col(colorblind_palette)
```

## STEP 3: Selecting and filtering data of interest
```{r Selecting and filtering data of interest }
data_curated <- data %>%
  select(record_id, privacy_acceptance___1, sex___1, sex___2, calculated_age, 
         lupus, sledai_points___1, sledai_points___2, sledai_points___3, sledai_points___4, sledai_points___5, sledai_points___6,
         sledai_points___7, sledai_points___8, sledai_points___9, sledai_points___10, sledai_points___11, sledai_points___12, sledai_points___13,
         sledai_points___14, sledai_points___15, sledai_points___16, sledai_points___17, sledai_points___18, sledai_points___19, 
         sledai_points___20, sledai_points___21, sledai_points___22, sledai_points___23, sledai_points___24, 
         aan, nephritis, prednisolona, immunesupresive, egfr,pregnancy, comorbidities, 
         health_provider, dx_time, time_symptoms, treatment___0, treatment___1, treatment___2, treatment___3, treatment___4,treatment___5,
         treatment___6, treatment___7, treatment___8, family_sle, home, ocupation, school,
         socioeconomic_status, cal_saludfisica, calc_psicologico, calc_relacionesociales, calc_ambiente, bref1, bref2) %>%
  filter(privacy_acceptance___1 == 1 &  (calculated_age >= 18 & calculated_age < 100)) %>%
  filter(sex___1 != sex___2 & sex___2 != sex___1) %>%
  filter(prednisolona != 7 | treatment___0 == 1 | treatment___1 == 1 | treatment___2 ==1 | treatment___3 ==1 | treatment___4 ==1 | treatment___5 ==1 | treatment___6 ==1 | treatment___7 ==1 | treatment___8 ==1 , preserve = TRUE) %>%
  mutate(lupus = replace(lupus, lupus == 2, "Ctrl")) %>%
  mutate(lupus = replace(lupus, lupus == 1, "Lupus")) %>%
  mutate(sex___1 = replace(sex___1, sex___1 == 1, "Women"))%>%
  mutate(sex___1 = replace(sex___1, sex___1 == 0, "Men")) %>%
  rename(sex=sex___1, Seizure=sledai_points___1, Psychosis=sledai_points___2, 'Organic brain syndome'=sledai_points___3, 'Visual disturbance'=sledai_points___4,
         'Cranial nerve disorder'=sledai_points___5, 'Lupus headache'=sledai_points___6, 'Cerebrovascular accident'=sledai_points___7,
         Vasculitis=sledai_points___8, Arthritis=sledai_points___9, Myositis=sledai_points___10, 'Urinary Casts'=sledai_points___11,
         Hematuria=sledai_points___12, Proteinuria=sledai_points___13, Pyuria=sledai_points___14, Rash=sledai_points___15,
         Alopecia=sledai_points___16, 'Mucosal ulcers'=sledai_points___17, 'Pleurisy'=sledai_points___18, Pericarditis=sledai_points___19,
         'Low complement'=sledai_points___20, 'Increased DNA binding'=sledai_points___21, Fever=sledai_points___22, Thrombocytopenia=sledai_points___23,Leukopenia=sledai_points___24,
         Corticosteoids=prednisolona, Antimalarials=treatment___0, Glucocorticoids=treatment___1, Mycophenolate=treatment___2, Azathioprine=treatment___3, Metothrexate=treatment___4,
         Rituximab=treatment___5, Cyclosporine=treatment___6, Cyclophosphamide=treatment___7, Others=treatment___8) %>%
  select(-record_id, -privacy_acceptance___1, -sex___2)
```

## STEP 4: Data selection for sociodemographic analysis
```{r Data selection }
sociodemographic_df <-  data_curated %>%
  select(sex, calculated_age, lupus,aan, nephritis,  immunesupresive, egfr,pregnancy, comorbidities, Corticosteoids,
         health_provider, dx_time, time_symptoms, Antimalarials, Glucocorticoids, Mycophenolate, Azathioprine, Metothrexate,
         Rituximab, Cyclosporine, Cyclophosphamide, Others, family_sle, ocupation, school) %>%  #To get all sociodemographic data
  filter(lupus == "Lupus") %>% #To only visualize people with SLE data
  mutate(health_provider = replace(health_provider, health_provider == 1, "IMSS")) %>%
  mutate(health_provider = replace(health_provider, health_provider == 2, "ISSSTE")) %>%
  mutate(health_provider = replace(health_provider, health_provider == 3, "PEMEX")) %>%
  mutate(health_provider = replace(health_provider, health_provider == 4, "SEDENA")) %>%
  mutate(health_provider = replace(health_provider, health_provider == 5, "SEMAR")) %>%
  mutate(health_provider = replace(health_provider, health_provider == 6, "SSa")) %>%
  mutate(health_provider = replace(health_provider, health_provider == 7, "SESA")) %>%
  mutate(health_provider = replace(health_provider, health_provider == 8, "IMSS-Bienestar")) %>%
  mutate(health_provider = replace(health_provider, health_provider == 9, "Private")) %>%
  mutate(pregnancy = replace(pregnancy, pregnancy == 0, "Pregnant")) %>%
  mutate(pregnancy = replace(pregnancy, pregnancy == 1, "Not pregnant")) %>%
  mutate(ocupation = replace(ocupation, ocupation == 1, "Student")) %>%
  mutate(ocupation = replace(ocupation, ocupation == 2, "Employee")) %>%
  mutate(ocupation = replace(ocupation, ocupation == 3, "Unemployed")) %>%
  mutate(ocupation = replace(ocupation, ocupation == 4, "Retired")) %>%
  mutate(school = replace(school, school == 1, "None")) %>%
  mutate(school = replace(school, school == 2, "Elementary school/9 years")) %>%
  mutate(school = replace(school, school == 3, "Junior high school/12 years")) %>%
  mutate(school = replace(school, school == 4, "Senior high school.Technician/15 years")) %>%
  mutate(school = replace(school, school == 5, "College")) %>%
  mutate(school = replace(school, school == 6, "Graduate school")) %>%
  mutate(nephritis = replace(nephritis, nephritis == 1, "Yes")) %>%
  mutate(nephritis = replace(nephritis, nephritis == 2, "No")) %>%
  mutate(aan = replace(aan, aan == 0, "Yes")) %>%
  mutate(aan = replace(aan, aan == 1, "No")) %>%
  mutate(comorbidities = replace(comorbidities, comorbidities == 1, "Diabetes mellitus")) %>%
  mutate(comorbidities = replace(comorbidities, comorbidities == 2, "Hypertension")) %>%
  mutate(comorbidities = replace(comorbidities, comorbidities == 3, "Cancer")) %>%
  mutate(comorbidities = replace(comorbidities, comorbidities == 4, "Cardiovascular disease")) %>%
  mutate(comorbidities = replace(comorbidities, comorbidities == 5, "Osteoporosis")) %>%
  mutate(comorbidities = replace(comorbidities, comorbidities == 8, "Thyroid diseases")) %>%
  mutate(comorbidities = replace(comorbidities, comorbidities == 9, "Other")) %>%
  mutate(comorbidities = replace(comorbidities, comorbidities == 11, "None"))
```

## STEP 5: Data selection for sociodemographic analysis
### *NA represents registries that were not filled for this questions*
```{r Percentages obtention }
### Percentage of participants by sex
sociodemographic_df %>%
  count(.$sex, "Women") %>% 
  mutate(percent = n/ sum(n) * 100)

### Mean age and mean age at diagnosis
## Age
mean_age <- mean(sociodemographic_df$calculated_age)
sd_age <- sd(sociodemographic_df$calculated_age)
mean_age
sd_age

## Duration of the disease
mean_dxtime <- mean(as.numeric(sociodemographic_df$dx_time))
sd_dxtime <- sd(sociodemographic_df$dx_time)
mean_dxtime
sd_dxtime

## Age at diagnosis
sociodemographic_df['age_at_dx'] <- sociodemographic_df$calculated_age - sociodemographic_df$dx_time # To #calculate age at diagnosis
mean_dxage <- mean(as.numeric(sociodemographic_df$age_at_dx))
mean_dxage
sd_dxage <- sd(sociodemographic_df$age_at_dx)
sd_dxage

### Percentage of participants by occupation
sociodemographic_df %>%
  count(.$ocupation, "0") %>% 
  mutate(percent = n/ sum(n) * 100)

### Percentage of participants by last academic degree
sociodemographic_df %>%
  count(.$school, "0") %>% 
  mutate(percent = n/ sum(n) * 100)

## Time since diagnosis
mean_dxtime <- mean(sociodemographic_df$dx_time)
sd_dxtime <- sd(sociodemographic_df$dx_time)
mean_dxtime
sd_dxtime

### Percentage of pregnant participants
sociodemographic_df %>%
  count(.$pregnancy , "Pregnant") %>% # 0 = pregnant
  mutate(percent = n/ sum(n) * 100)

### Percentage of participants with a familiar with SLE
sociodemographic_df %>%
  count(.$family_sle , 0) %>% # 0 = si
  mutate(percent = n/ sum(n) * 100)

### Percentage of participants with nephritis
sociodemographic_df %>%
  count(.$nephritis, 0) %>% # 0 = si
  mutate(percent = n/ sum(n) * 100)

### Percentage of participants with comorbidities
sociodemographic_df %>%
  count(.$comorbidities, 1) %>% 
  mutate(percent = n/ sum(n) * 100)

### Percentage of health providers 
sociodemographic_df %>%
  count(.$health_provider , 0) %>% 
  mutate(percent = n/ sum(n) * 100)
 
### Percentage of participants with comorbidities
data_curated %>%
  count(.$Corticosteoids, 1) %>% 
  mutate(percent = n/ sum(n) * 100)

##Percentage of participants by treatment
##Antimalarials
sociodemographic_df %>%
  count(.$Antimalarials, 1) %>%  # 1=They consume the treatment
  mutate(percent = n/ sum(n) * 100)

##Glucocorticoids
sociodemographic_df %>%
  count(.$Glucocorticoids, 1) %>% # 1=They consume the treatment
  mutate(percent = n/ sum(n) * 100)

##Corticoids
sociodemographic_df %>%
  count(.$Corticosteoids, 0) %>% # 1=They consume the treatment
  mutate(percent = n/ sum(n) * 100)

##Mycophenolate
sociodemographic_df %>%
  count(.$Mycophenolate, 1) %>% # 1=They consume the treatment
  mutate(percent = n/ sum(n) * 100)

##Azathioprine
sociodemographic_df %>%
  count(.$Azathioprine, 1) %>% # 1=They consume the treatment
  mutate(percent = n/ sum(n) * 100)

##Metothrexate
sociodemographic_df %>%
  count(.$Metothrexate, 1) %>% # 1=They consume the treatment
  mutate(percent = n/ sum(n) * 100)

#Rituximab
sociodemographic_df %>%
  count(.$Rituximab, 1) %>% # 1=They consume the treatment
  mutate(percent = n/ sum(n) * 100)

#Cyclosporine
sociodemographic_df %>%
  count(.$Cyclosporine, 1) %>% # 1=They consume the treatment
  mutate(percent = n/ sum(n) * 100)

#Cyclophosphamide
sociodemographic_df %>%
  count(.$Cyclophosphamide, 1) %>% # 1=They consume the treatment
  mutate(percent = n/ sum(n) * 100)

#Others
sociodemographic_df %>%
  count(.$Others, 1) %>% # 1=They consume the treatment
  mutate(percent = n/ sum(n) * 100)
```

## STEP 6: Age and time since diagnosis distribution
### Age distribution
```{r Age distribution }
cal_age <- data_curated %>%  
  filter(lupus=="Lupus") %>%
  count(calculated_age)

mean_age <- mean(data_curated$calculated_age)
sd(data_curated$calculated_age)

### Distribution of patients according to age
ggplot(cal_age, (mapping = aes(x = factor(calculated_age), y = n))) + 
  geom_col(width = 0.5, color="#6bb099", fill="#6bb099") + 
  geom_text(aes(label = n), vjust = -0.5, size=3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3)) + #orientation text in bars
  xlab("Age") + 
  ylab("# of patients") +  
  ggtitle("Age of patients") + 
  scale_fill_manual(values = colorblind_palette) +
  theme(plot.title = element_text(hjust = 0.5))  

### Distribution of patients according to groups by age
## Groups by age
cal_age <- data_curated %>%  
  filter(lupus=="Lupus") %>%
  count(calculated_age, sex)

cal_age <-cal_age %>%
  mutate(group = (if_else(calculated_age <= 20, "18-20 yrs", "21-30 yrs"))) %>%
  mutate(group = replace(group, calculated_age>=31 ,"31-40 yrs")) %>%
  mutate(group = replace(group, calculated_age>=41,"41-50 yrs")) %>%
  mutate(group = replace(group, calculated_age>50,"51-99 yrs"))

#Plot based on number of patients (women/men)
ggplot(cal_age, aes(x = group, y = n, fill = sex)) + 
  geom_col(position = position_dodge(0.5), width = 0.5) + 
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0.3, size=11)) +
  xlab("Age group") + 
  ylab("Number of patients")  + 
  ggtitle("Patients by age group") + 
  theme(plot.title = element_text(hjust = 0.5)) +  
  scale_fill_manual(values = colorblind_palette) +
  labs(fill = "")

group_sex <- cal_age %>%
  count(group, sex, wt = n) 

#Total patients by group 
valores_totales <- data.frame(group=
                                c("18-20 yrs", "21-30 yrs", "31-40 yrs", "41-50 yrs", "51-99 yrs"), 
                              sex___1=
                                c("Total", "Total", "Total", "Total", "Total"), 
                              n=
                                c(group_sex$n[1], sum(group_sex$n[2:3]), 
                                  sum(group_sex$n[4:5]),  sum(group_sex$n[6:7]), 
                                  sum(group_sex$n[8:9])))

# Collapsing 2 df in 2
valores_totales <- valores_totales %>%
  rename(sex=sex___1)
group_sex <- rbind(group_sex, valores_totales)

#Plot based on percentage of patients
group_sex <-group_sex %>%
  mutate(percentage = round(((n*100)/colSums(valores_totales[3])), 2))
group_sex$sex <- factor(group_sex$sex, levels=c("Women", "Men", "Total"))

ggplot(group_sex, (mapping = aes(x = factor(group), y = percentage, fill=factor(sex)))) + 
  geom_col(position = position_dodge(0.567), width = 0.5) +
  scale_fill_manual(values = colorblind_palette) +
  labs(y = "Percentage (%)", 
       fill = "",
       x = "Groups by age",
       title = "Patients distribution according to age")  +
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 8),
        axis.title = element_text(size = 12, face = "bold"),
        axis.title.x = element_text(hjust = 0.5),
        axis.text.y = element_text(size =10),
        legend.position = "bottom")
```
### Time since diagnosis distribution
```{r Time since diagnosis distribution }
cal_dx_time <- data_curated %>%  
  count(dx_time, sex)
cal_dx_time <- na.omit(cal_dx_time)

ggplot(cal_dx_time, (mapping = aes(x = factor(dx_time), y = n))) + 
  geom_col(width = 0.5, color="#6bb099", fill="#6bb099")+ 
  scale_fill_manual(values = colorblind_palette) +
  geom_text(aes(label = n), vjust = -0.5, size=3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3)) +
  xlab("Time since dx (yrs)") + 
  ylab("Number of patients") +   
  ggtitle("Time since diagnosis") + 
  theme(plot.title = element_text(hjust = 0.5), 
        panel.grid.major = element_blank(),
        panel.background= element_rect(fill = "transparent", colour = "black"), # To let the axis of the graph 
        plot.background = element_rect(fill = "transparent", colour = NA)) # To remove the background

cal_dx_time  <- cal_dx_time  %>%
  mutate(group = (if_else(dx_time <= 1, "1 year or less", "1-5 yrs"))) %>%
  mutate(group = replace(group, dx_time>=5 ,"6-10 yrs")) %>%
  mutate(group = replace(group, dx_time>=11,"11-25 yrs")) %>%
  mutate(group = replace(group, dx_time>26,"26-50 yrs")) %>%
  mutate(group = replace(group, dx_time>51,"51 yrs or more")) 

ggplot(cal_dx_time, aes(x = factor(group), y = n, fill = factor(sex))) + 
  geom_col(position = position_dodge(0.5), width = 0.5) + 
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0.3, size=11)) +
  xlab("Time since diagnosis yrs") + 
  ylab("Number of patients")  + 
  ggtitle("Patients by group of time of diagnosis") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_fill_manual(values = colorblind_palette) +
  labs(fill = "")

group_dxtime_sex <- cal_dx_time %>%
  count(group, sex, wt = n) 

total_timedx <- data.frame(group=
                             c("1 year or less", "1-5 yrs", "6-10 yrs", "11-25 yrs", "26-50 yrs", "51 yrs or more"),
                           sex=
                             c("Total", "Total", "Total", "Total", "Total", "Total"),
                           n=
                             c(sum(group_dxtime_sex$n[1:2]), sum(group_dxtime_sex$n[3:4]), 
                               sum(group_dxtime_sex$n[9:10]),  sum(group_dxtime_sex$n[5:6]), 
                               sum(group_dxtime_sex$n[7:8]), sum(group_dxtime_sex$n[9])))
total_timedx <- na.omit(total_timedx)
group_dxtime_sex <- rbind(group_dxtime_sex, total_timedx)

group_dxtime_sex <-group_dxtime_sex %>%
  mutate(percentage = round(((n*100)/colSums(total_timedx[3])), 2))
group_dxtime_sex$sex <- factor(group_dxtime_sex$sex, levels=c("Women", "Men", "Total"))

ggplot(group_dxtime_sex, (mapping = aes(x = fct_relevel(group, c("1 year or less", "1-5 yrs", "6-10 yrs", "11-25 yrs", "26-50 yrs")), 
                                        y = percentage, fill=factor(sex)))) + 
  geom_col(position = position_dodge(0.567), width = 0.5) +
  scale_fill_manual(values = colorblind_palette) +
  labs(y = "Percentage (%)", 
       fill = "",
       x = "Groups according to disease duration",
       title = "Patients grouped by disease duration")  +
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5, size = 8),
      axis.title = element_text(size = 12, face = "bold"),
      axis.title.x = element_text(hjust = 0.5),
      axis.text.y = element_text(size =10),
      legend.position = "bottom")
```
### Age at diagnosis distribution
```{r Age at diagnosis distribution }
agedx <- sociodemographic_df %>%  
  count(age_at_dx, sex)

agedx <-agedx %>%
  mutate(group = (if_else(age_at_dx <= 18, "< 18 yrs", "18-20 yrs"))) %>%
  mutate(group = replace(group, age_at_dx>=21 ,"21-30 yrs")) %>%
  mutate(group = replace(group, age_at_dx>=31 ,"31-40 yrs")) %>%
  mutate(group = replace(group, age_at_dx>=41,"41-50 yrs")) %>%
  mutate(group = replace(group, age_at_dx>50,"51-99 yrs"))

#Plot based on number of patients (women/men)
ggplot(agedx, aes(x = group, y = n, fill = sex)) + 
  geom_col(position = position_dodge(0.5), width = 0.5) + 
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0.3, size=11)) +
  xlab("Age group") + 
  ylab("Number of patients")  + 
  ggtitle("Patients by age at diagnosis") + 
  theme(plot.title = element_text(hjust = 0.5)) +  
  scale_fill_manual(values = colorblind_palette) +
  labs(fill = "")

group_dxage_sex <- agedx %>%
  count(group, sex, wt = n) 

#Total patients by group 
valores_totalesagedx <- data.frame(group=
                                c("< 18 yrs", "18-20 yrs", "21-30 yrs", "31-40 yrs", "41-50 yrs", "51-99 yrs"), 
                              sex=
                                c("Total", "Total", "Total", "Total", "Total", "Total"), 
                              n=
                                c(group_dxage_sex$n[1:2], sum(group_dxage_sex$n[3:4]), 
                                  sum(group_dxage_sex$n[5:6]),  sum(group_dxage_sex$n[7:8]), 
                                  sum(group_dxage_sex$n[9:10])))

# Collapsing 2 df in 2
group_dxage_sex <- rbind(group_dxage_sex, valores_totalesagedx)

#Plot based on percentage of patients
group_dxage_sex <-group_dxage_sex %>%
  mutate(percentage = round(((n*100)/colSums(valores_totalesagedx[3])), 2))
group_dxage_sex$sex <- factor(group_dxage_sex$sex, levels=c("Women", "Men", "Total"))

ggplot(group_dxage_sex, (mapping = aes(x = factor(group), y = percentage, fill=factor(sex)))) + 
  geom_col(position = position_dodge(0.567), width = 0.5) +
  scale_fill_manual(values = colorblind_palette) +
  labs(y = "Percentage (%)", 
       fill = "",
       x = "Groups by age at diagnosis",
       title = "Patients distribution according to age at diagnosis")  +
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 8),
        axis.title = element_text(size = 12, face = "bold"),
        axis.title.x = element_text(hjust = 0.5),
        axis.text.y = element_text(size =10),
        legend.position = "bottom")
```

## STEP 7: Demographic distribution
```{r Demographic distribution }
#### If mxmaps is not installed applied the next code
#if (!require(devtools)) {
 # install.packages("devtools")
#}
#devtools::install_github('diegovalle/mxmaps')
#install.packages('mxmaps')

#### Data preparation 
States <- data_curated %>% 
  mutate(home = replace(home, home == 1, "Aguascalientes")) %>%
  mutate(home = replace(home, home == 2, "Baja California")) %>%
  mutate(home = replace(home, home == 3, "Baja California Sur")) %>%
  mutate(home = replace(home, home == 4, "Campeche")) %>%
  mutate(home = replace(home, home == 5, "Coahuila")) %>%
  mutate(home = replace(home, home == 6, "Colima")) %>%
  mutate(home = replace(home, home == 7, "Chiapas")) %>%
  mutate(home = replace(home, home == 8, "Chihuahua")) %>%
  mutate(home = replace(home, home == 9, "Durango")) %>%
  mutate(home = replace(home, home == 10, "Ciudad de México")) %>%
  mutate(home = replace(home, home == 11, "Guanajuato")) %>%
  mutate(home = replace(home, home == 12, "Guerrero")) %>%
  mutate(home = replace(home, home == 13, "Hidalgo")) %>%
  mutate(home = replace(home, home == 14, "Jalisco")) %>%
  mutate(home = replace(home, home == 15, "México")) %>%
  mutate(home = replace(home, home == 16, "Michoacán")) %>%
  mutate(home = replace(home, home == 17, "Morelos")) %>%
  mutate(home = replace(home, home == 18, "Nayarit")) %>%
  mutate(home = replace(home, home == 19, "Nuevo León")) %>%
  mutate(home = replace(home, home == 20, "Oaxaca")) %>%
  mutate(home = replace(home, home == 21, "Puebla")) %>%
  mutate(home = replace(home, home == 22, "Querétaro")) %>%
  mutate(home = replace(home, home == 23, "Quintana Roo")) %>%
  mutate(home = replace(home, home == 24, "San Luis Potosí")) %>%
  mutate(home = replace(home, home == 25, "Sinaloa")) %>%
  mutate(home = replace(home, home == 26, "Sonora")) %>%
  mutate(home = replace(home, home == 27, "Tabasco")) %>%
  mutate(home = replace(home, home == 28, "Tamaulipas")) %>%
  mutate(home = replace(home, home == 29, "Tlaxcala")) %>%
  mutate(home = replace(home, home == 30, "Veracruz")) %>%
  mutate(home = replace(home, home == 31, "Yucatán")) %>%
  mutate(home = replace(home, home == 32, "Zacatecas")) %>%
  filter(lupus == "Lupus") %>%
  count(home) 

regions <- data.frame(
  "region" = 1:32,
  "state_name" = c("Aguascalientes", "Baja California", "Baja California Sur", "Campeche", "Coahuila", "Colima", "Chiapas", "Chihuahua",
  "Ciudad de México", "Durango", "Guanajuato", "Guerrero","Hidalgo", "Jalisco", "México", "Michoacán", "Morelos", "Nayarit", "Nuevo León",
  "Oaxaca", "Puebla", "Querétaro", "Quintana Roo", "San Luis Potosí", "Sinaloa", "Sonora", "Tabasco", "Tamaulipas", "Tlaxcala", "Veracruz",
  "Yucatán", "Zacatecas")) # Order of states according to mxmaps dispositions

home <- merge(States, regions, by.x = "home", by.y = "state_name")
home <- home %>%
  rename(registers=n, state_name=home)

### Mapping registries
breaks <- c(0, 11, 26, 51, 76, 101, 150, Inf) 
colors <- c("#fde0dd","#fcc5c0","#fa9fb5","#f768a1","#dd3497",
"#ae017e","#7a0177")
home$breaks <- cut(home$registers, breaks)
home$value <- home$breaks 

##Graph map
mxstate_choropleth(home,
                   title = "Geographic distribution of people \n with SLE in México", legend = "Number of patients")+ 
  scale_fill_manual(breaks = levels(home$breaks),
                    values = colors,
                    labels=c("1-10", "11-25", "26-50", "51-75", "76-100", "101-150", "151-175"))+
  labs(fill="Number of \n patients")+
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 8),
        legend.position = "right",
        legend.title = element_text(size=14),
        legend.text = element_text(size = 12, hjust = 0.5),
        legend.key.size = unit(0.5, 'cm'),
        legend.title.align = 0.5)
```

## STEP 8: Socioeconomic status
```{r Calculating socioeconomic status}
#### Selection of the data
s_status <- data_curated%>%
  filter(lupus=="Lupus")%>%
  count(socioeconomic_status)

#### Assignation of socioeconomic level
s_status <-s_status %>%
  mutate(group = (if_else(socioeconomic_status <= 47, "E = 261 USD", "D = 430 USD"))) %>%
  mutate(group = replace(group, socioeconomic_status>=94 & socioeconomic_status<=115,"D+ = 595 USD")) %>%
  mutate(group = replace(group, socioeconomic_status>=116 & socioeconomic_status<=140,"C- = 754 USD")) %>%
  mutate(group = replace(group, socioeconomic_status>=141 & socioeconomic_status<=167,"C = 962 USD")) %>%
  mutate(group = replace(group, socioeconomic_status>=168 & socioeconomic_status<=201,"C+")) %>%
  mutate(group = replace(group, socioeconomic_status>=202,"A/B"))

#### Summarizing patients by socioeconomic level
se_status <- data.frame(group=
                               c("E = 261 USD", "D = 430 USD", "D+ = 595 USD", "C- = 754 USD", "C = 962 USD"),
                           n=
                             c(sum(s_status$n[1:20]), sum(s_status$n[21:52]), 
                               sum(s_status$n[53:69]),  sum(s_status$n[70:77]), 
                               sum(s_status$n[78:79])))
se_status[is.na(se_status)] <-0
  
#### Obtaining percentages
se_status <- se_status %>%
  mutate(percentage = round(((n*100)/colSums(se_status[2])), 1)) %>%
  rename(Category = group)

#### Making the plot
se_status$Category = factor(se_status$Category, levels = c("C = 962 USD", "C- = 754 USD", "D+ = 595 USD",
                                                           "D = 430 USD", "E = 261 USD" ))

ggplot(se_status,aes(x =Category, y = percentage, fill= Category)) +
  geom_col(position = position_dodge(0.567), width = 0.5)+
  scale_fill_manual(values = c("E = 261 USD"= "#fdd49e", 
                               "D = 430 USD"= "#fdbb84",
                               "D+ = 595 USD" ="#fc8d59",
                               "C- = 754 USD" ="#e34a33",
                               "C = 962 USD" = "#b30000")) +
  labs(y = "Percentage (%)", 
       fill = "Category",
       x = "Socioeconomic levels",
       title = "Socioeconomic level distribution")  +
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        legend.position = "right",
        legend.title = element_text(size=12),
        legend.text = element_text(size = 9),
        axis.title = element_text(size = 10, face = "bold"),
        legend.key.size = unit(1, 'cm'),
        legend.title.align = 0.5)
```

## STEP 8: Clinical data
### Comorbidities
```{r Analyzing comorbidities}
## Identifying the variables
comorbidities  <- data_curated %>%
  filter(lupus=="Lupus")%>%
  mutate(comorbidities = replace(comorbidities, comorbidities == 1, "Diabetes mellitus")) %>%
  mutate(comorbidities = replace(comorbidities, comorbidities == 2, "Hypertension")) %>%
  mutate(comorbidities = replace(comorbidities, comorbidities == 3, "Cancer")) %>%
  mutate(comorbidities = replace(comorbidities, comorbidities == 4, "Cardiovascular disease")) %>%
  mutate(comorbidities = replace(comorbidities, comorbidities == 5, "Osteoporosis")) %>%
  mutate(comorbidities = replace(comorbidities, comorbidities == 8, "Thyroid diseases")) %>%
  mutate(comorbidities = replace(comorbidities, comorbidities == 9, "Other")) %>%
  mutate(comorbidities = replace(comorbidities, comorbidities == 11, "None")) # %>%

lupus_total <- comorbidities %>%
  count(lupus)
lupus_total
comorbidities <- comorbidities %>%
  count(comorbidities, lupus)

## Eliminate options associated with other autoimmune diseases
comorbidities <- comorbidities[!(comorbidities$comorbidities == '6'),]
comorbidities <- comorbidities[!(comorbidities$comorbidities == '0'),]
comorbidities <- comorbidities[!(comorbidities$comorbidities == '7'),]

## Obtaining percentages
comorbidities <- comorbidities %>%
  mutate(percentage = round(((n*100)/(lupus_total$n)))) %>%
  arrange(desc(percentage))

## Making the plot
ggplot(comorbidities, aes(x=reorder(comorbidities, +percentage), y = percentage, fill=""))+
  geom_bar(stat = "identity")+
  coord_flip()+
  scale_fill_manual(values = "#2b8cbe") +
  labs(y = "Percentage (%)", 
       fill = "",
       x = "Comorbidity",
       title = "Comorbidities \n in SLE patients")  +
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        legend.position = "none",
        axis.title = element_text(size = 12, face = "bold"),
        axis.text.y = element_text(size =12),
        axis.text.x = element_text(size=12))
```

### Clinical manifestations
```{r Analyzing clinical manifestations}
## Selecting data 
clinical_df <- data_curated %>%
  filter(lupus=="Lupus") %>%
  select( "Seizure", "Psychosis", "Organic brain syndome", "Visual disturbance", "Cranial nerve disorder","Lupus headache",
          "Cerebrovascular accident","Vasculitis","Arthritis", "Myositis",  "Urinary Casts",  "Hematuria", "Proteinuria",
          "Pyuria", "Rash", "Alopecia", "Mucosal ulcers", "Pleurisy", "Pericarditis","Low complement","Increased DNA binding",
          "Fever","Thrombocytopenia","Leukopenia") 

## Principal manifestations
## Data preparation
sumcolumn_sledai <- colSums(clinical_df)
sumcolumn_sledai

clinical_prop <- data.frame(
  "Manifestation" = c("Seizure", "Psychosis", "Organic brain syndrome", "Visual disturbance", "Cranial nerve disorder","Lupus headache",
                      "Cerebrovascular accident","Vasculitis","Arthritis", "Myositis",  "Urinary Casts",  "Hematuria", "Proteinuria",
                      "Pyuria", "Rash", "Alopecia", "Mucosal ulcers", "Pleurisy", "Pericarditis","Low complement","Increased DNA binding",
                      "Fever","Thrombocytopenia","Leukopenia"),
  "n" = sumcolumn_sledai[c(1:24)]
)
clinical_prop <- clinical_prop %>%
  mutate(percentage = round(((n*100)/lupus_total$n))) 

## Making the plot
ggplot(clinical_prop, aes(x=reorder(Manifestation, +percentage), y = percentage, fill=""))+
  geom_bar(stat = "identity")+
  coord_flip()+
  scale_fill_manual(values = "#800026") +
  labs(y = "Percentage (%)", 
       fill = "",
       x = "Clinical manifestation",
       title = "Principal clinical manifestations \n in SLE patients")+
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.9, size = 14, face = "bold"),
        legend.position = "none",
        axis.title = element_text(size = 12, face = "bold"),
        axis.title.x = element_text(hjust = 0.5),
        axis.text.y = element_text(size =10),
        axis.text.x = element_text(size =10))

## Visualizing intersections frequency between different severe manifestations.
## Selecting only severe manifestations (according to SLEDAI criteria)
severe_manif <- clinical_df %>%
  select(-Rash, -Alopecia, -'Mucosal ulcers', -Pleurisy, -Pericarditis,
  -'Low complement', -'Increased DNA binding', -Fever, -Thrombocytopenia, -Leukopenia)

severe_manif <- as.data.frame(sapply(severe_manif , as.numeric))
str(severe_manif)

## Making the plot
upset(severe_manif, nsets = 14, text.scale = c(1.5, 1.25, 1.25, 1, 1.5, 1.5),
      sets.bar.color = c("#882255", "#888888","#882255", "#888888", "#888888", "#888888",
                         "#882255", "#882255", "#888888", "#888888", "#882255","#888888",
                         "#888888", "#882255"), 
      mb.ratio = c(0.50, 0.5),
      mainbar.y.label = "Frequency (n)", sets.x.label = "Total frequency (n)")
```