---
title: "03 SLE Quality of life and economic stratum"
author: "Domingo Martínez"
date: "2023-01-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this script we will try to answer the following question:

*total quality of life changes in SLE-individuals across different socioeconomic levels?*

Now we only consider SLE individual. In this analysis we will take into consideration the variables listed following:

- cal_totalcv: aggregated quality of life, a score greater than 40 indicates a high quality of life.

- socioeconomic_status: aggregate economic level/status.

## STEP 1: Reading database
```{r Reading database}
library(tidyverse)
tb.pacientes <- read_csv("DATA_2023-01-05_completas/DATA_2023-01-05_0147_Complete_db_patients_under_SLE_medication_5.csv")
```

## STEP 2: Selecting variables of interest
```{r Picking variables of interest}
tb.pacientes <- tb.pacientes %>%
  select(lupus,
         sex___1, sex___2,
         calculated_age,
         bref1, bref2,
         calc_ambiente,
         cal_saludfisica,
         calc_psicologico,
         calc_relacionesociales,
          cal_totalcv,
         socioeconomic_status
         )
```

## STEP 3: Checking how many NAs we had
<https://cran.r-project.org/web/packages/naniar/vignettes/naniar-visualisation.html>
```{r Checking NAs}
sum(is.na(tb.pacientes))
# There are 1917 missing values
library(naniar)
vis_miss(tb.pacientes)
tb.pacientes<-na.omit(tb.pacientes)
```

Finally, we get 966 SLE patients with completed registry.

## STEP 4: Renaming variables
```{r Renaming variables}
library(dplyr)
tb.pacientes <- tb.pacientes %>%
  rename("Lupus" = lupus,
         "Female" = sex___1,
         "Male" = sex___2,
         "Calculated Age" = calculated_age,
         "Quality of life perception" = bref1,
         "Health perception" = bref2,
         "Environmental QoL-domain" = calc_ambiente,
         "Physical health QoL-domain" = cal_saludfisica,
         "Psychologial QoL-domain" = calc_psicologico,
         "Social relations QoL-domain" = calc_relacionesociales,
         "Total quality of life" = cal_totalcv,
         "Socioeconomic status" = socioeconomic_status)
```


## STEP 5: Formating class of variables
```{r Formating class of variables}
str(tb.pacientes)
#library(tidyverse)
tb.pacientes <- tb.pacientes %>%
  mutate(Lupus = as_factor(Lupus),
         Female = as_factor(Female),
         Male = as_factor(Male)) 

str(tb.pacientes)
```

```{r Re-ordering factors}
tb.pacientes$Female <- tb.pacientes$Female %>%
  recode_factor("1" = "1", "0" = "0")

tb.pacientes$Male <- tb.pacientes$Male %>%
  recode_factor("1" = "1", "0" = "0")

str(tb.pacientes)
```
## STEP 6 Checking female-male proportion
```{r}
table(tb.pacientes[,2:3])
```

The SLE patients groups comprise 966 participants, 906 of them registered as women so, we have next proportions:

0.9378882 Women.
0.0621118 Men.

## STEP 7: Recoding socioeconomic status to categorical level, according to AMAI.

According to AMAI, *the evaluation of socioeconomic status* contemplates 7 levels:
     A/B (202+ points)
     C+ (168-201) 
     C (141-167)
     C- (116-140)
     D+ (95-115)
     D (48-94) and 
     E (0-47). 
     
A higher score implies a better capacity to fulfill the necessities within a household, next we will recode the points to categorical level.

```{r Calculating socioeconomic levels}
# We use cut() function to put breaks and get labels. And save the result into new column called "Socioeconomic level"
tb.pacientes$`Socioeconomic level` <- cut(tb.pacientes$`Socioeconomic status`,
    breaks = c(0,47,94,115,140,167,201,1000),
    labels = c("E","D","D+","C-","C","C+","A/B"),
    right = T)
str(tb.pacientes)
# Let's see how many participants belong to each level.
table(tb.pacientes$`Socioeconomic level`)
```
## STEP 8: Summary of data
```{r}
summary(tb.pacientes)
```
**IMPORTANT:** We had atypical ages so, we need to control such variable, also we had sex imbalance, then we need to control such variable too.

## STEP 9: The quality of life in SLE patients change among economic stratum?

Aplying Kruskal.Wallis test by socioeconomic level to test differences in quality of life, specifically, ´Total quality of life´
```{r Testing total quality of life among socioeconomic levels}
kruskal.test(tb.pacientes$`Total quality of life`,
             tb.pacientes$`Socioeconomic level`)
```

Yes, there are differences of quality of life depending on the socioeconomic leve, now we have to find where such differences are.

```{r Finding differences of quality of life by pairwise permutation test}
library(rcompanion)
TQoL <- tb.pacientes$`Total quality of life`
SocioeconomicLevel <- as.character(tb.pacientes$`Socioeconomic level`)
str(SocioeconomicLevel)
resultados.pairwise <-pairwisePermutationTest( x=tb.pacientes$`Total quality of life`, 
                         g=SocioeconomicLevel,
                         method = "fdr")
resultados.pairwise
resultados.pairwise$p.adjust.expanded <- format(resultados.pairwise$p.adjust,scientific = F)
resultados.pairwise
```
**INTERPRETATION:** 
There is no difference between C and C-
C is over D by 4.079 points in TQoL scale.
There is only marginal difference between C and D+
C is over E by 4.358 points in TQoL scale.
C- is over D by 4.476 in TQoL scale.
Between C- and D+ there is no difference.
C- is higher than E by 4.285 points in the TQoL scale.
D+ is higher than D by 5.884 points in the TQoL scale.
There is only a marginal differences between D and E.
D+ is higher than E by 4.819 points in the TQoL scale.

## STEP 10 Ploting such comparison

<https://stackoverflow.com/questions/17084566/put-stars-on-ggplot-barplots-and-boxplots-to-indicate-the-level-of-significanc>

<https://r-graph-gallery.com/269-ggplot2-boxplot-with-average-value.html>

```{r Ploting}
# Library
library(ggplot2)
library(ggsignif)

# plot
p <- ggplot(tb.pacientes, aes(x=`Socioeconomic level`, y=`Total quality of life`, fill=`Socioeconomic level`)) +
    geom_boxplot(alpha=0.5) +
  ylim(15,140) +
    stat_summary(fun=mean, geom="point", shape=5, size=2.5, color="black") +
    scale_fill_brewer(palette="RdPu") +
  theme_bw() +
  geom_signif(comparisons=list(c("D+", "E")), annotations="***", y_position = 90, size = 0.2) +
  geom_signif(comparisons=list(c("D+", "D")), annotations="***", y_position = 80, size = 0.2) +
  geom_signif(comparisons=list(c("C-", "E")), annotations="***", y_position = 100, size = 0.2) +
  geom_signif(comparisons=list(c("C-", "D")), annotations="***", y_position = 110, size = 0.2) +
  geom_signif(comparisons=list(c("C", "E")), annotations="***", y_position = 120, size = 0.2) +
  geom_signif(comparisons=list(c("C", "D")), annotations="***", y_position = 130, size = 0.2) +
  ggtitle("Quality of life by economic stratum in SLE patients")
p

ggsave("SLE_QoL_and_economic_stratum.png", units = c("cm"), dpi = 300 )
```







