---
title: "Binding all three databases"
author: "Domingo Martínez"
date: "2023-01-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this script we will bind all three databases.

In this integration we will take into consideration the variables listed following:

- lupus: 1 indicates a participant diagnosed with LES.
         2 indicates a participant NON-diagnosed with LES.  

- sex: 1 or sex___1 indicates female participant.
       2 or sex___2 indicates male participant.

- calculated_age: age calculated automatically based on birthday.

- bref1: quality of life perception, how would you rate your QoL?

- bref2: healt perception, how satisfied are you with your health?

- cal_saludfisica:  physical domain of quality of life (QoL).

- calc_psicologico: psychological domain of QoL.

- calc_relacionesociales: social domain of QoL.

- calc_ambiente: environmental domain od QoL.

- cal_totalcv: aggregated quality of life, a score greater than 40 indicates a high quality of life.

## STEP 1: Reading all three databases
```{r reading databases}
library(tidyverse)
tb.SLE.patients <- read.csv("DATA_2023-01-05_completas/DATA_2023-01-05_0147_Complete_db_patients_under_SLE_medication_5.csv")
tb.RedCap.volunters <- read.csv("DATA_2023-01-05_completas/DATA_2023-01-05_0147_Complete_db_controls_from_RedCap.csv")  
tb.Twins.volunters <- read.csv("DATA_2023-01-05_completas/DATA_2023-01-13_twin_control_lupus.csv")
```

## STEP 2: Picking and renaming variables of interest

```{r Picking formating and renaming variables of interes SLE patients}
tb.SLE.patients<-tb.SLE.patients %>%
  select(lupus,
         sex___1, sex___2,
         calculated_age,
         bref1, bref2,
         calc_ambiente,
         cal_saludfisica,
         calc_psicologico,
         calc_relacionesociales,
          cal_totalcv) %>%
  rename("Lupus" = lupus,
         "Female" = sex___1,
         "Male" = sex___2,
         "Calculated Age" = calculated_age,
         "Quality of life perception" = bref1,
         "Health perception" = bref2,
         "Environmental QoL-domain" = calc_ambiente,
         "Physical health QoL-domain" = cal_saludfisica,
         "Psychologial QoL-domain" = calc_psicologico,
         "Social relations QoL-domain" = calc_relacionesociales,
         "Total quality of life" = cal_totalcv)
## Formating variables
tb.SLE.patients<-tb.SLE.patients %>%
  mutate(Lupus = as_factor(Lupus),
         Female = as_factor(Female),
         Male = as_factor(Male))
## Reordering factors
  tb.SLE.patients$Female<-tb.SLE.patients$Female %>%
  recode_factor("1" = "1", "0" = "0")

tb.SLE.patients$Male <- tb.SLE.patients$Male %>%
  recode_factor("1" = "1", "0" = "0")
## Aggregate group label
tb.SLE.patients$Group <- rep("SLEpatients",length(tb.SLE.patients$Lupus))
str(tb.SLE.patients)
```

```{r Picking formating and renaming variables of interes RedCap volunters}
tb.RedCap.volunters<-tb.RedCap.volunters%>%
  select(lupus,
         sex___1, sex___2,
         calculated_age,
         bref1, bref2,
         calc_ambiente,
         cal_saludfisica,
         calc_psicologico,
         calc_relacionesociales,
          cal_totalcv) %>%
  rename("Lupus" = lupus,
         "Female" = sex___1,
         "Male" = sex___2,
         "Calculated Age" = calculated_age,
         "Quality of life perception" = bref1,
         "Health perception" = bref2,
         "Environmental QoL-domain" = calc_ambiente,
         "Physical health QoL-domain" = cal_saludfisica,
         "Psychologial QoL-domain" = calc_psicologico,
         "Social relations QoL-domain" = calc_relacionesociales,
         "Total quality of life" = cal_totalcv)
## Formating variables
tb.RedCap.volunters<-tb.RedCap.volunters %>%
  mutate(Lupus = as_factor(Lupus),
         Female = as_factor(Female),
         Male = as_factor(Male))
## Reordering factors
  tb.RedCap.volunters$Female<-tb.RedCap.volunters$Female%>%
  recode_factor("1" = "1", "0" = "0")

tb.RedCap.volunters$Male<-tb.RedCap.volunters$Male %>%
  recode_factor("1" = "1", "0" = "0")
## Aggregate group label
tb.RedCap.volunters$Group<-rep("RedCapVolunter", length(tb.RedCap.volunters$Lupus))
str(tb.SLE.patients)
```

```{r Picking formating and renaming variables of interes Twins volunters}
## Formating class
tb.Twins.volunters$sexo <- as.factor(tb.Twins.volunters$sexo)
## Now, let's calculate # sex___1 and sex___2 variables
tb.Twins.volunters$sex___1 <-tb.Twins.volunters$sexo %>%
  recode_factor("M" = "1", "H" = "0")
tb.Twins.volunters$sex___2 <-tb.Twins.volunters$sexo %>%
  recode_factor("M" = "1", "H" = "0")
## Calculating lupus variabl
tb.Twins.volunters$Lupus <- rep("2", length(tb.Twins.volunters$record_id))
tb.Twins.volunters$Lupus<- as.factor(tb.Twins.volunters$Lupus)
## Renaming variables
tb.Twins.volunters<-tb.Twins.volunters %>%
  select(Lupus,
         sex___1,
         sex___2,
         edad,
         qol_bref1,
         qol_bref2,
         qol_ambiente,
         qol_saludfisica,
         qol_psicologico,
         qol_relacionessociales,
         qol_total) %>%
  rename("Female" = sex___1,
         "Male" = sex___2,
         "Calculated Age" = edad,
         "Quality of life perception" = qol_bref1,
         "Health perception" = qol_bref2,
         "Environmental QoL-domain" = qol_ambiente,
         "Physical health QoL-domain" = qol_saludfisica,
         "Psychologial QoL-domain" = qol_psicologico,
         "Social relations QoL-domain" = qol_relacionessociales,
         "Total quality of life" = qol_total)
## Adding group label
tb.Twins.volunters$Group <- rep("TwinVolunter", length(tb.Twins.volunters$Female))
```

## STEP 3: Binding all three datasets
```{r binding all three datasets}
tb.completo<-bind_rows(tb.SLE.patients,
          tb.RedCap.volunters,
          tb.Twins.volunters)

tb.completo$Group <-  as.factor(tb.completo$Group)

str(tb.completo)
write.csv(tb.completo, "DATA_2023-01-05_completas/base.de.datos.integrada.csv")
```

## STEP 4: Checking data quality
```{r Visualazing NA}
sum(is.na(tb.completo))
# There are 7387 missing values
library(naniar)
vis_miss(tb.completo)
```

```{r Visualazing pairs plot}
library(GGally)
ggpairs(tb.completo, aes(colour=Group, alpha=0.75))
ggsave("pair.plot.base.de.datos.integrada.png", plot = last_plot(), width = 60, height = 55, units = c("cm"), device = png, dpi = 300)
```

NOTES:

There is a participant approximately 1200 years old. We need to impute or delete that observation.

In SLE patients, age only correlated with environmental domain of QoL.



