---
title: "Curating duplicated registries"
author: "Domingo Martínez"
date: "2023-01-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Inthis sript we will find duplicated registries into the RedCap Lupus Project Production <https://redcap.lavis.unam.mx/apps/redcap/redcap_v10.9.2/DataExport/index.php?pid=88>, once we identified that duplicated registries, it will be necessary to handly verify which of them is the most complete depending on the purpose of the studies.

**IMPORTANT:** When the script finished, the original database was eliminated to avoid misuse of the data.

## Reading data base
```{r Readig database}
library(dplyr)
library(forcats)
tb.LupusRGMX<-read.csv("LupusProjectProducti_DATA_2023-01-08_1704.csv")
glimpse(tb.LupusRGMX)
```
## Counting how many times each cellphone is repeated
```{r Cellphone number cross tabulation}
# Creating a cross table for cellphone numbers
cellphones <- tb.LupusRGMX %>%
  group_by(cell_number) %>%
  tally() %>%
  na.omit() %>%
  mutate(n=as.factor(n)) %>%
  glimpse()
# Let's see how many times each registry appear.
  levels(cellphones$n)
# Now we will eliminate those appearing once.
cellphones <- cellphones %>%
  filter(n!=1)
```

## Match repeated cellphones to ID registries
```{r Matching duplicated cellphone to record_id}
cellphones.vector <- cellphones$cell_number

tb.results.1 <- tibble(record_id = numeric(),
                       cell_number = numeric())

for (i in cellphones.vector) {
    result<-tb.LupusRGMX %>%
    select(record_id, cell_number) %>%
    filter(cell_number == i)
tb.results.1 <- rbind(tb.results.1,result)
}
write.csv(tb.results.1, "duplicate_registries_01.csv")
```

## Counting how many times each email is repeated
```{r Cellphone number cross tabulation}
# Creating a cross table for emails
emails <- tb.LupusRGMX %>%
  group_by(email) %>%
  tally() %>%
  filter(email!="") %>%
  mutate(n=as.factor(n)) %>%
  glimpse()
# Let's see how many times each registry appear.
  levels(emails$n)
# Now we will eliminate those appearing once.
emails <- emails %>%
  filter(n!=1)
```

## Match repeated emails to ID registries
```{r Matching duplicated email to record_id}
emails.vector <- emails$email

tb.results.2 <- tibble(record_id = numeric(),
                       cell_number = numeric())

for (i in emails.vector) {
    results<-tb.LupusRGMX %>%
    select(record_id, email) %>%
    filter(email == i)
tb.results.2 <- rbind(tb.results.2,results)
}
write.csv(tb.results.2, "duplicate_registries_02.csv")
```

```{bash delete data set}
cd ~/Documentos/LupusRGMX/Curating_RedCap_database/
rm -rf LupusProjectProducti_DATA_2023-01-08_1704.csv
```

## 
```{r read duplicated-registries inform}
library(dplyr)
library(forcats)
# file.choose()
tb.cuestionarios.complete<-read.csv("/home/domingo/Documentos/PosDoc_Liigh_Lupus/Analisis_tratamiento&calidad_de_vida/Curating_RedCap_database/LupusProjectProducti-Cuestionarioscomplet_DATA_2023-01-25_1721.csv")
# Select variables of interest
tb.cuestionarios.complete<-tb.cuestionarios.complete[ , -c(2,3,4,5,7) ]
write.csv(tb.cuestionarios.complete, "informe_cuestionarios_completos.csv")
```

```{bash delete bd}
cd ~/Documentos/LupusRGMX/Curating_RedCap_database/
rm- rf LupusProjectProducti-Cuestionarioscomplet_DATA_2023-01-25_1721.csv
```

## From duplicated registries pick the most complete

From the report titled ""informe_cuestionarios_completos.csv" we select those duplicated registries and will pick up the most complete registry, based on "Complete?" variables.

The codes for this variables are:
0	Incomplete
1	Unverified
2	Complete

```{r reading both datasets}
tb.duplicated.registries <- read.csv("duplicate_registries_02.csv")
tb.completed.questionnaires <- read.csv("informe_cuestionarios_completos.csv")
```

Get a unique list for duplicated mails
```{r list of duplicated mails}
list.of.duplicated<- unique(tb.duplicated.registries$email)
str(list.of.duplicated)
table(list.of.duplicated)
```

Iterate over duplicated-list to compare duplicated registries to select the most completed registries. Later, these registries can be removed from list
```{r picking the most completed from duplicated registries}
library(dplyr)

mejores.id.de.los.repetidos <-tibble(id.mejor = as.numeric())

for (i in 1:length(list.of.duplicated)) {
  correo <-list.of.duplicated[i]

  a <- tb.completed.questionnaires %>%
    filter(email == correo)
  
  suma_completo <- apply(a[,-c(1,2,3)], 1,sum)
  
  record_id <- a$record_id 
   
  c <- tibble(record_id, suma_completo)
  
  d <- c[order(-suma_completo),]
  id.best  <- d[1,1] # This record_is is the most completes from the n-mail repeted
 
  mejores.id.de.los.repetidos <- rbind(mejores.id.de.los.repetidos,id.best) 
    
}

```

## Now let's create a list of worst registries

That's mean, from the repeated registries, let's take off the best registry and kept the worst registries in a black list. These registries will be drooped off from the data set of LupusRGMX.

From tb.duplicated.registries let's take record_id column
```{r making a list of trash registries}
es.el.registro.mas.completo <-ids.repetidos$record_id %in% mejores.id.de.los.repetidos$record_id

tb.es.el.registro.mas.completo<-tibble(ids.repetidos, es.el.registro.mas.completo)

lista.registros.duplicados.para.eliminar <- tb.es.el.registro.mas.completo %>%
  filter(es.el.registro.mas.completo == F)

write.csv(lista.registros.duplicados.para.eliminar, "lista.registros.duplicados.para.eliminar.csv")
```

```{bash removing databases with mails}
cd ~/Documentos/LupusRGMX/Curating_RedCap_database/
rm- rf duplicate_registries_01.csv duplicate_registries_02.csv informe_cuestionarios_completos.csv 

```

