---
title: "Filtering cases of interest"
author: "Domingo Martínez"
date: "2023-01-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here, we will filter the complete databases from RedCap attending two criterions:

1. Drop out duplicated registries.
2. From SLE patients, selecting those that already take medical treatment.

## STEP 1: Droping out duplicated registries.

We will eliminate those duplicated registries showed in the file titled:*lista.registros.duplicados.para.eliminar*, this is a black list containing all duplicated registries have to be eliminated.

```{r reading black list}
library(dplyr)
black.list <- read.csv("lista.registros.duplicados.para.eliminar.csv")
```

```{r reading RedCap database}
bd <- read.csv("LupusProjectProducti_DATA_2023-01-05_0147.csv")
```

```{r filtering duplicated registries}
library(Hmisc)
bd <- subset(bd, record_id %nin% black.list$record_id)
```

## STEP 2: Selecting SLE patients under medical treatment

- The variable "lupus" indicate 1, if the participant has been SLE diagnosed.
- The variable "prednisolona" indicate if participant takes corticosteroids, specifically:

0	Prednisona
1	Prednisolona
2	Deflazacort
3	Meticorten
4	Metilprednisolona
5	Calcort
6	Betametazona
7	No
8	Otro

Then, we will take those participants diagnosed with SLE (lupus == 1), and also we will exclude those participants don't taking any corticosteroid or taking other medication (i.e, "Otro".)

- The variable “treatment” indicate what medications patient takes on a regular basis to treat SLE.

0	treatment___0	Antimaláricos (Hidroxicloroquina, cloroquina, plaquenil)
1	treatment___1	Corticoesteroides (Prednisona, prednisolona, deflazacort, meticorten, metilprednisolona, calcort, betametazona)
2	treatment___2	Ácido micofenólico (Micofenolato, cellcept, mofetil)
3	treatment___3	Azatioprina (imuran, azatrilem)
4	treatment___4	Metotrexato (Ledertrexate)
5	treatment___5	Rituximab
6	treatment___6	Ciclosporina
7	treatment___7	Ciclofosfamida
8	treatment___8	Otros

all these "treatment___n" variables indicate yes with 1, and not with 0.

Thus, we also will select those taking any of those medications. We exclude those answering "Otros" (i.e. other medications) because we can't be sure that medicine is for treating SLE.

Briefly, selection criteria are:

- lupus == 1
- prednisolona %in% c(0,1,2,3,4,5,6)
- treatment %in% c(0,1,2,3,4,5,6,7,8)

```{r picking SLE-diagnosed and taking corticosteroids or SLE treatment}
bd1<-bd %>%
  filter(lupus == 1) %>%
  filter(prednisolona %in% c(0,1,2,3,4,5,6) | 
           treatment___0 == 1 |
           treatment___1 == 1 |
           treatment___2 == 1 |
           treatment___3 == 1 |
           treatment___4 == 1 |
           treatment___5 == 1 |
           treatment___6 == 1 |
           treatment___7 == 1 ) 
bd1
```


## Exporting complete db of participants taking any SLE treatment or medication
```{r exporting complete SLE db}
write.csv(bd1, "DATA_2023-01-05_0147_Complete_db_patients_under_SLE_medication_5.csv")
```

## Picking and exporting controls database from RedCap
```{r exporting controls database from RedCap}
bd2<-bd %>%
  filter(lupus == 2)
write.csv(bd2, "DATA_2023-01-05_0147_Complete_db_controls_from_RedCap.csv")
```




